### 2022-08-10 main/chapter_2ljets.tex
modelling proofread

\subsubsection{Alternative \texorpdfstring{$\zjets$}{Z+jets}}
\label{sec:2ljets_zjets_alt}
As mentioned in Section~\ref{sec:2ljets_low}, the $\zjets$ sample comparison
is hugely important to our results in SR-Low.
Its central yields are displayed in Figure~\ref{fig:2ljets_low_sr_or_cr_region}
in the union of \crz\ and SR-Low, which form a continuous region in $\mjj$.
Our central $\zjets$ samples use \sherpa~2.2.1~\cite{
Bothmann:2019yzt,
Gleisberg:2008ta,
ATL-PHYS-PUB-2016-003
}
at next-to-leading order for up to $4$ partonic jets, and leading order for
$5$ partonic jets.
In Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt} we compare this to our
alternative sample that simulates its hard scatter
with \madgraph\ (\amcatnlo~2.2.3~\cite{Alwall:2014hca}) at leading order with
up to four partonic jets, plugged into \pythia~8~\cite{Sjostrand:2014zea} for
showering and hadronization.
This \madgraph\ alternative sample simulates only $\mll > 40\,\eV[G]$, so in
regions allowing $\mll$ outside of this we evaluate relative variations and
scale them up to the central yield.
The plots in Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt} show statistical
error bars only; the alternative sample has larger statistical errors since
it contains around one sixth as many simulated events as the central sample,
and that difference is not countered by their different weight distributions
in this region.

Since \sherpa\ implements both the hard scatter and shower models, we are
considering variations due to the simultaneous change of both of these
components together.

For their implementation in the background model in HistFactory, these
variations reduce to the relative scale factors that are presented in
Table~\ref{tab:2ljets_zjets_alt}.
Note that up and down variations of the sample comparison parameter do not
mean up and down variations of bin yields, but to the coupled variations
in directions given by the signs of the relative variations given in the
various regions.
These variations are evaluated according to a standard recipe prescribed by
the ATLAS SUSY group~\cite{atlas_twiki_susytheoretical} that says to
\begin{itemize}
\item evaluate the difference between the two samples $\delta_{12}$.
\item evaluate the statistical uncertainty on the alternative sample $\sigma_2$.
\item combine these components in quadrature, so
$\sigma_{12} = \sqrt{\delta_{12}^2 + \sigma_2^2}$, and
\item if the deviation is statistically significant, then preserve the direction
of the uncertainty, so $\sigma = \mathrm{sgn}(\delta_{12})\sigma_{12}$.
\end{itemize}
An `if A then B' statement does not define what to do if A is false, so
this recipe does not define what sign to give $\sigma$ if the deviation is not
statistically significant.
Since all variations have direction
--- they are either positive or negative or zero ---
we are unable to give no direction to any deviation, and we choose to
use the same, signed, definition of $\sigma$ for all regions.

As seen from Table~\ref{tab:2ljets_zjets_alt}, the statistical uncertainty
$\sigma_2$ is a substantial contribution to these variations, and so
the statistical uncertainty in the alternative sample does significantly
expand this variation.
Indeed, $\sigma_2$ is the larger contribution in SR-Low-a and SR-Low-b, which
are the regions most sensitive to $\zjets$ variations.
Our unprescribed choice to match their directions therefore matters.
It may have been reasonable to treat them differently, perhaps by adding
independent parameters in the separate regions.
I believe such a change would have improved the plausibility of our background
modelling, and I regret not finding the capacity to push for it at the time.

@@

Including this $\zjets$ sample comparison greatly improves our model's post-fit
agreement with the SR-Low deficits.
It therefore seems proper to disclose that this variation was added after
unblinding, and after we all knew about those deficits.
Prior to unblinding, we in the $\twoljets$ team had not planned on including it
in addition to other theoretical uncertainties, and I do not know whether or
not committee decisions in the collaboration would have required its inclusion
without the deficit.
The \atlas\ SUSY recommendation is to consider the comparison if
``the scale ($+$ PDF) uncertainties are too small in your
analysis''~\cite{atlas_twiki_susytheoretical},
but I know of no official advice on how to judge what is ``too small''.
Unblinding procedures are discussed further in
Section~\ref{sec:2ljets_unblinding}.

@@

Auxiliary data to the $\twoljets$ paper~\cite{atlas2022searches} include
Figure~\ref{fig:2ljets_low_crz_pre_post_post}~\cite{hepdata.116034.v1/t101},
which appears to show good agreement in SR-Low and \crz, a relatively
small error bar in SR-Low of around $\pm10$ events per $100\,\eV[G]$,
and a curious dip in the $\zjets$ background in the signal region.
The figure caption also admits that
``All EWK search control and signal regions are included in the fit.''

Fitting only to control region (not signal regions) results in
Figure~\ref{fig:2ljets_low_crz_pre_post_pre}, which shows a clear deficit in
the signal region, with \diboson\ yields peaking near to $m(W)$ above a smooth
varying $\zjets$ background.
Fitting to the signal region slashes the $\zjets$ background in SR-Low by
around $80\%$. How?
This sample comparison is the largest variation in VR-Low, and we have seen
from Table~\ref{tab:2ljets_zjets_alt} that SR-Low variations are around $30\%$.
A few factors contribute:
\begin{itemize}
\item The $\zjets$ sample is normalized to \crz\ with an unconstrained
normalization factor. This means that variations to the $\zjets$ sample in \crz
are absorbed by scaling all of its modifications in other regions.
\item As listed in Table~\ref{tab:2ljets_zjets_alt}, positive variations
in \crz\ match to negative variations in SR-Low bins; the two yields are
anticorrelated.
With normalization to \crz, this anticorrelation increases the effective
variation in SR-Low to $\pm40\%$.
\item Post-fit, the parameter controlling this $\zjets$ sample comparison
modifier is pulled to around $2\sigma$. This is seen in
Figure~\ref{fig:2ljets_fit_parameters}, and will be discussed further in
Section~\ref{sec:2ljets_results}.
\end{itemize}
A two-sigma pull on a $40\%$ scaling therefore explains this $80\%$ change.
Validation regions do not constrain our fits, and are also affected by
this effect; in VR-Low, small differences between the central and alternative
$\zjets$ samples are visible in Figure~\ref{fig:2ljets_vrlow_alt}, but
after fitting to the signal regions,
Figure~\ref{fig:2ljets_low_vrlow_pre_post} shows that the $\zjets$ sample
is almost wiped out, leaving a deficit below data.

@@

It is not clear that such a large extrapolation from the difference between
two simulations is physically reasonable, especially since a large fraction
of its effect is due to the statistical error on a small alternative sample.
To avoid this questionable effect, I searched for alternative configurations
of the background model, such as constraining the parameter to $1$-sigma.
Committee decisions within the \atlas\ SUSY group dictated that we should keep
it as described, reasoning along the lines that
a) this is a real and significant deficit in SR-Low, so must appear as either
a pull on systematic variations or as a deviation from data, and
b) we prefer the pull on systematic variations.

With this decision, post-fit results (such as the summary plot of
Figure~\ref{fig:2ljets_summary})
indeed show no significant deviation from data in SR-Low-a or SR-Low-b,
and this $\zjets$ sample comparison is the main reason why.
Internal reviewers were, however, confused as to why we mentioned a deficit in
SR-Low that did not appear in the results plots or tables.
Much to my ire, the paper editors therefore removed all mentions of an SR-Low
deficit from the paper.
Despite bring real and significant, this deficit was erased from public view,
and to my knowledge, is only reported through inner details of
the HistFactory models shared in our \hepdata\ archive~\cite{hepdata.116034}.

To my eyes, Figure~\ref{fig:2ljets_low_crz_pre_post} suggests that the
excessive \diboson\ background should also take some blame for this poor
prediction in SR-Low;
its \diboson\ yields peak in the signal region on $\mjj \approx m(W)$, and
we do not expect large $W\rightarrow jj$ contributions in this \diboson\ process,
since one boson is needed for the leptons and the other is needed for the
$\met$ that we select.
We shall see in Section~\ref{sec:2ljets_diboson_alt} why \diboson\ yields are
not similarly squashed by sample comparison variations.

\subsubsection{Alternative \diboson}
\label{sec:2ljets_diboson_alt}
No alternative samples were available for comparison with our central \diboson
samples, which are simulated with \sherpa~2.2.2 at NLO with up to six
electroweak vertices.
This limitation is simply a result of how \atlas\ has decided to allocate
its limited resources on, and quite reasonably so, since full simulation is
computationally expensive.
We did, however, have access to the cheaper alternative set of small,
truth-level samples, that were simulated with the same \sherpa\ setup but with
varied parameters to assess uncertainties in the modelling of parton showers.

@@

In the signal regions, most of these statistical variations are around
$100\%$ relative to the central yield, with a minimum across all variations
of $26\%$ and a maximum of $960\%$.
For $Q_\textrm{cut}^\textrm{CKKW}$ and $S_Q$ their signs appear randomly
distributed in positive and negative directions in approximately equal
proportions, but for \texttt{CSS\_KIN} most are correlated in the positive
direction, with the only SR-OffShell-a moving in the negative direction, by
$46\%$.
Variations in control regions are typically smaller, ranging from $12\%$ to
$50\%$, but remain limited by statistical precision.

To include these large variations moving in wacky directions could allow the
background model to easily fit deviations on the basis of our ignorance
from statistical imprecision.
To mitigate this, we instead take the variations from looser regions
VR-sys-main and VR-sys-OffShell, that target on- and off-shell regions
respectively and hope to improve the statistical precision.
Both require the pre-selection from Table~\ref{tab:2ljets_presel};
VR-sys-main also requires
$\mjj \in 60\textrm{--}110\,\eV[G]$ and $\nbtag \leq 1$,
and VR-sys-OffShell requires
$\nbtag = 0$, $\ptjone > 100\,\eV[G]$ and $\dphijmet > 2$ to target the
OffShell regions.
No extra lepton selections are made for these regions since these variations
relate primarily to jets through parton showers, and not to leptons, and
relaxing cuts can only improve statistical precision.

@@

\paragraph{Alternative $\ttbar$}
Several $\ttbar$ variations model alternative simulations of the
hard scatter (central \powheg\ versus \amcatnlo),
parton shower (central \pythia8 versus \herwig7).
Initial state radiation variations are evaluated from an alternative
sample with $h_\textrm{damp} = 517.5\,\eV[G] \approx 3m(t)$,
which damps real emissions with weights to vary an ``eigentune'' parameter
named \texttt{Var3c}~\cite{ATL-PHYS-PUB-2014-021}..
The central sample uses a default value $h_\textrm{damp} = 172.5\,\eV[G]$ to
approximate $m(t)$~\cite{CMS-PAS-TOP-16-021}.

@@

Scale variations mostly move all regions in the same direction, with increased
scales tending to reduce yields.
Variations in $\mu_R$ are typically in the $\approx 10\textrm{--}30\%$ range,
making them larger than the $\approx 1\textrm{--}10\%$ variations due to
$\mu_F$.
Given that it takes a maximum, the envelope prescription gives a single larger
variation, but that may not mean more flexibility in the model since the SUSY
prescription has three degrees of freedom.

@@

\paragraph{Parton distribution variations}
Like scale variations, uncertainties in parton distribution functions (PDFs)
are assigned by weights stored Fuks:2012qxby the generator.
These weights, however, do not correspond to simple parameters variations,
but different replica PDFs in a Monte Carlo ensemble
that is intended to model the uncertainties left after fitting to data
in the NNPDF methodology~\cite{NNPDF:2017mvq,BALL2011112,BALL2012608}.

The number of PDF replicas varies between samples, but is between
$100$ and $138$ for all except $\zyjets$, that has just one central and two
alternatives.
Remarkably, both SUSY and PMG prescriptions agree on how to assign variations
from these replica weights, since both say to follow the PDF4LHC
recommendations~\cite{PDF4LHC_Butterworth_2016} (specifically, p.
Since PDF4LHC gives two alternative prescriptions, one using mean and variance
and the other using quantiles, we evaluate and consider both.
Both are negligible, making $\approx 1\%$ differences before normalization
to control regions, but are included for completeness.
Since there is nothing particularly special about the central PDF weight,
this prescription allows the central scale-factor to differ from unity,
and this is also a $\approx 1\%$ effect.

@@

\subsection{Signal variations}
\label{sec:2ljets_signal_variations}
Signal uncertainties are a strange case.
Like Standard Model backgrounds, our signal models are simulated by
approximations to their ideal QFT, and those approximations do have
uncertainties that we can describe.
On backgrounds, however, large uncertainties are conservative, since
flexibility to fit outlying data reduces the chance of mistaken claims of
discoveries.
That effect reverses for signals.
In maximum \heplikelihood\ fits, a more flexible
signal model will always fit any observed data better, and in the limit of
infinite flexibility, an unconstrained signal will always fit
any data better than a constrained background, making it look better to certain
interpretations.%
\footnote{%
This superiority of flexible models is not fundamental, but is a property of
\heplikelihood\ fitting methods.
If we want to punish models that make uncertain predictions, then we should
consider likelihoods (not \emph{maximum} \heplikelihood s), as described in
Section~\ref{sec:searches_searches}.
The important distinction is between uncertain predictions from the model,
and our uncertainty in what the model prediction is.
Our ignorance of a model's predictions should not lead us to excluding it.
}

@@

\subsubsection{Theory variations}
\label{sec:2ljets_theory_variations}
Another signal variation is assigned to cover sources of uncertainty from
scale variations ($\mu_R$, $\mu_F$, and the merging scale
$Q_\textrm{cut}^\textrm{CKKW}$)) and modelling of radiations and parton
showers (`eigentune' parameters
\texttt{Var1},
\texttt{Var2},
\texttt{Var3a},
\texttt{Var3b},
\texttt{Var3c}~\cite{ATL-PHYS-PUB-2014-021}).
Following the \atlas\ SUSY group
prescription~\cite{atlas_twiki_susytheoretical},
we consider only a coherent scaling of the scales $\mu_R$ and $\mu_F$ together,
so this theory variation includes a total of seven components.

Scale variations have appeared here twice:
first for cross-section variations and
second for these theory variations.
They therefore contribute both in each fit model and in the varied
cross-sections for varied exclusion contours.
This is questionable, but we shall see here that, thankfully, this scale
double-counting makes very little difference.

\paragraph{Extrapolation}
To evaluate each variation requires a new simulation with different parameters,
so to reduce computational costs we use truth-level only
(through SimpleAnalysis~\cite{simpleanalysis_cern}) and evaluate the
variations at only a handful of benchmark signal models.
From these benchmark signal models, we evaluate a relative uncertainty due
due to each component, and extrapolate those variations from these benchmark
examples to others in the parameter plane.

Extrapolation is performed as follows.
Each benchmark point $i$ has a location in the parameter plane $\vec y_i$, and
for a given source of uncertainty $j$ its relative variation has been evaluated
as $r_{ij}$.
To extrapolate these variations across each parameter plane to another signal
model at $\vec x$, we use an exponentially weighted combination
\begin{equation}
\label{eqn:2ljets_signal_extrap}
r_j(\vec x) = \frac{
\sum_i r_{ij} e^{-d(\vec x, \vec y_i)}
}{
\sum_i e^{-d(\vec x, \vec y_i)}
}
\end{equation}
in which $d(\vec x, \vec y)$ is a distance function chosen to encode some
physical understanding into the extrapolation,
and the sum is over all benchmark points.

@@

Relative uncertainty $r$ is evaluated for each component, and the total
uncertainty is their sum in quadrature, capped at $100\%$.
This is done independently in each selection region without considering the
directions of coherent variations; all variations are combined as described
and varied together in all regions under the control of a single parameter in
the fit.

\begin{figure}[tp]
\centering
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_signal_sys_c1n2_SRHigh4.pdf}
\caption{C1N2, SR-High-1J}
\end{subfigure}
\hfill
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_signal_sys_c1n2_SROffShell_1.pdf}
\caption{C1N2, SR-OffShell-a}
\end{subfigure}
\\[0.5em]
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_signal_sys_gmsb_SRHigh8_2.pdf}
\caption{GMSB, SR-High-1J}
\end{subfigure}
\hfill
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_signal_sys_gmsb_SRllbb.pdf}
\caption{GMSB, SR-OffShell-a}
\end{subfigure}
\caption[
Example extrapolated signal theoretical uncertainties
]{%
Example extrapolated signal theoretical uncertainties.
Variations are evaluated for signals at the marked points and extrapolated to
each parameter plane according to Equation~\ref{eqn:2ljets_signal_extrap}.
To account for its drastically different kinematics, the C1N2 off-shell space
is treated independently.
}
\label{fig:2ljets_signal_sys_extrap}
\end{figure}

\paragraph{Evaluation}
Statistical precision has limited the utility of most variations described in
this chapter.
These are no different.
To reduce statistical noise here, I generated truth samples $10$~times as large
as our central, reconstructed samples.
These sample sizes vary roughly with cross-section from $300$k to
$2.8$m events per sample, and due to the $\sqrt{n}$ scaling of
Monte Carlo errors, means approximately $\sqrt{10}$ times reduced
standard deviation.
However, this remains insufficient since each resulting variation is a
sum in quadrature of variations in $7$ pairs of alternative samples
whose combination will inflate the uncertainty by roughly a factor of
$\sqrt{7}$.
Where signal models' yields has significant in a signal region,
each variation tends to be around $\approx 20\%$ relative error,
and are compatible with the central samples within statistical uncertainties.

>>>>

\subsubsection{Alternative \texorpdfstring{$\zjets$}{Z+jets}}
\label{sec:2ljets_zjets_alt}
As seen in Section~\ref{sec:2ljets_low}, the $\zjets$ sample comparison is
important to our results in SR-Low.
Its central yields are displayed in Figure~\ref{fig:2ljets_low_sr_or_cr_region}
in the union of \crz\ and SR-Low, which form a continuous region in $\mjj$.
Our central $\zjets$ samples use \sherpa~2.2.1~\cite{
Bothmann:2019yzt,
Gleisberg:2008ta,
ATL-PHYS-PUB-2016-003
}
at next-to-leading order for up to $4$ partonic jets, and leading order for
$5$ partonic jets.
In Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt} we compare this to our
alternative sample that simulates its hard scatter
with \madgraph\ (\amcatnlo~2.2.3~\cite{Alwall:2014hca}) at leading order with
up to four partonic jets, plugged into \pythia~8~\cite{Sjostrand:2014zea} for
showering and hadronization.
This \madgraph\ alternative sample simulates only $\mll > 40\,\eV[G]$, so in
regions allowing $\mll$ outside of this we evaluate relative variations and
scale them up to the central yield.
The plots in Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt} show statistical
error bars only; the alternative sample has larger statistical errors since
it contains around one sixth as many simulated events as the central sample,
and that difference is not countered by their different weight distributions
in this region.

Since \sherpa\ implements both the hard scatter and parton shower models, by
comparing these two samples we are considering variations due to the
simultaneous change of both of hard scatter and parton shower components
together.

For their implementation in the background model in HistFactory, these
variations reduce to the relative scale factors that are presented in
Table~\ref{tab:2ljets_zjets_alt}.
Note that up and down variations of the sample comparison parameter do not
necessarily mean up and down variations of bin yields, but to the coupled
variations in directions given by the signs of the relative variations given in
the various regions.
These variations are evaluated according to a standard recipe prescribed by
the ATLAS SUSY group~\cite{atlas_twiki_susytheoretical} that says to
\begin{itemize}
\item evaluate the difference between the two samples $\delta_{12}$.
\item evaluate the statistical uncertainty on the alternative sample $\sigma_2$.
\item combine these components in quadrature, so
$\sigma_{12} = \sqrt{\delta_{12}^2 + \sigma_2^2}$, and
\item if the deviation is statistically significant, then preserve the direction
of the uncertainty, so $\sigma = \mathrm{sgn}(\delta_{12})\sigma_{12}$.
\end{itemize}
An `if A then B' statement does not define what to do if A is false, so
this recipe does not define what sign to give $\sigma$ if the deviation is not
statistically significant.
Since all variations have direction
--- they are either positive or negative or zero ---
we are unable to give no direction to any deviation, and we choose to
use the same, signed, definition of $\sigma$ for all regions.

As seen from Table~\ref{tab:2ljets_zjets_alt}, the statistical uncertainty
$\sigma_2$ is a substantial contribution to these variations, so
the statistical uncertainty in the alternative sample does significantly
expand this variation.
Indeed, $\sigma_2$ is the larger contribution in SR-Low-a and SR-Low-b, which
are the regions most sensitive to $\zjets$ variations.
Our unprescribed choice to match their directions therefore matters.
It may have been reasonable to treat them differently, perhaps by adding
independent parameters to describe the statistical errors in the separate
regions.
Such a change may well have improved the plausibility of our background
modelling, and I regret not finding the capacity to push for it at the time.

@@

Including this $\zjets$ sample comparison greatly improves our model's post-fit
agreement with the SR-Low deficits.
It therefore seems proper to disclose that this variation was added after
unblinding, and after we all knew about those deficits.
Prior to unblinding, we in the $\twoljets$ team had not planned on including it
in addition to other theoretical uncertainties, and I do not know whether or
not committee decisions in the collaboration would have required its inclusion
without the deficit.
The \atlas\ SUSY recommendation is to consider the comparison if
``the scale ($+$ PDF) uncertainties are too small in your
analysis''~\cite{atlas_twiki_susytheoretical},
but I know of no official advice on how to judge what is ``too small''.
Our unblinding procedure is discussed further in
Section~\ref{sec:2ljets_unblinding}.

@@

Auxiliary data to the $\twoljets$ paper~\cite{atlas2022searches} include
Figure~\ref{fig:2ljets_low_crz_pre_post_post}~\cite{hepdata.116034.v1/t101},
which appears to show good agreement in SR-Low and \crz, a relatively
small error bar in SR-Low of around $\pm10$ events per $100\,\eV[G]$,
and a curious dip in the $\zjets$ background in the signal region.
The figure caption also admits that
``All EWK search control and signal regions are included in the fit.''
Fitting only to control region (not signal regions) results in
Figure~\ref{fig:2ljets_low_crz_pre_post_pre}, which shows a clear deficit in
the signal region, with \diboson\ yields peaking near to $m(W)$ above a smooth
varying $\zjets$ background.
Fitting to the signal region slashes the $\zjets$ background in SR-Low by
around $80\%$. How?
This sample comparison is the largest variation in VR-Low, and we have seen
from Table~\ref{tab:2ljets_zjets_alt} that SR-Low variations are around $30\%$.
A few factors contribute:
\begin{itemize}
\item The $\zjets$ sample is normalized to \crz\ with an unconstrained
normalization factor. This means that variations to the $\zjets$ sample in \crz
are absorbed by scaling all of its modifications in other regions.
\item As listed in Table~\ref{tab:2ljets_zjets_alt}, positive variations
in \crz\ match to negative variations in SR-Low bins; the two yields are
anticorrelated.
With normalization to \crz, this anticorrelation increases the effective
variation in SR-Low to $\pm40\%$.
\item Post-fit, the parameter controlling this $\zjets$ sample comparison
modifier is pulled to around $2\sigma$. This is seen in
Figure~\ref{fig:2ljets_fit_parameters}, and will be discussed further in
Section~\ref{sec:2ljets_results}.
\end{itemize}
A two-sigma pull on a $40\%$ scaling therefore explains this $80\%$ change.
Validation regions do not constrain our fits, and are also affected by
this effect; in VR-Low, small differences between the central and alternative
$\zjets$ samples are visible in Figure~\ref{fig:2ljets_vrlow_alt}, but
after fitting to the signal regions,
Figure~\ref{fig:2ljets_low_vrlow_pre_post} shows that the $\zjets$ sample
is almost wiped out, leaving an apparent post-fit excess of data in VR-Low.

@@

It is not clear that such a large extrapolation from the difference between
two simulations is physically reasonable, especially since a large fraction
of its effect is due to the statistical error on a small alternative sample.
To avoid this questionable effect, I explored alternative configurations
of the background model, such as constraining the parameter to $1$-sigma.
Committee decisions within the \atlas\ SUSY group stated that we should keep
it as described, reasoning along the lines that
a) this is a real and significant deficit in SR-Low, so must appear as either
a pull on systematic variations or as a deviation from data, and
b) we prefer the pull on systematic variations.

With this decision, post-fit results (such as the summary plot of
Figure~\ref{fig:2ljets_summary})
indeed show no significant deviation from data in SR-Low-a or SR-Low-b,
and this $\zjets$ sample comparison is the main reason why.
Internal reviewers were, there, confused as to why we mentioned a deficit in
SR-Low that did not appear in the results plots or tables.
Much to my ire, the paper editors therefore removed all mentions of an SR-Low
deficit from the paper.
Despite bring real and significant, this deficit was erased from public view,
and to my knowledge, is only reported through inner details of
the HistFactory models shared in our \hepdata\ archive~\cite{hepdata.116034}
(and now, I hope, this thesis).

To my eyes, Figure~\ref{fig:2ljets_low_crz_pre_post} suggests that the
excessive \diboson\ background should also take some blame for this poor
prediction in SR-Low;
its \diboson\ yields peak in the signal region on $\mjj \approx m(W)$, and
we do not expect large $W\rightarrow jj$ contributions in this \diboson\
process since one boson is needed for the leptons and the other is needed for
the $\met$ that we select.
We shall see in Section~\ref{sec:2ljets_diboson_alt} why \diboson\ yields are
not similarly squashed by sample comparison variations.

\subsubsection{Alternative \diboson}
\label{sec:2ljets_diboson_alt}
No complete alternative samples were available for comparison with our central
\diboson\ samples, which are simulated with \sherpa~2.2.2 at NLO with up to six
electroweak vertices.
This limitation is simply a result of how \atlas\ has decided to allocate
its limited resources on, and quite reasonably so since full simulation is
computationally expensive and \diboson\ processes are relatively rare.
We did, however, have access to the cheaper alternative set of small,
truth-level samples, that were simulated with the same \sherpa\ setup but with
varied parameters to assess uncertainties in the modelling of parton showers.

@@

In the signal regions, most of these statistical variations are around
$100\%$ relative to the central yield, with a minimum across all variations
of $26\%$ and a maximum of $960\%$.
For $Q_\textrm{cut}^\textrm{CKKW}$ and $S_Q$ their signs appear randomly
distributed in positive and negative directions in approximately equal
proportions, but for \texttt{CSS\_KIN} most are correlated in the positive
direction, with the only SR-OffShell-a moving in the negative direction by
$46\%$.
Variations in control regions are typically smaller, ranging from $12\%$ to
$50\%$, but remain limited by statistical precision.

To include these large variations moving in wacky directions could allow the
background model to easily fit deviations on the basis of our ignorance
from statistical imprecision.
To mitigate this, we instead take the variations from looser regions
VR-sys-main and VR-sys-OffShell, that target on- and off-shell regions
respectively and hope to improve the statistical precision.
Both of these loose VR-sys-* regions require the pre-selection from
Table~\ref{tab:2ljets_presel};
VR-sys-main also requires
$\mjj \in 60\textrm{--}110\,\eV[G]$ and $\nbtag \leq 1$,
and VR-sys-OffShell requires
$\nbtag = 0$, $\ptjone > 100\,\eV[G]$ and $\dphijmet > 2$ to target the
OffShell regions.
No extra lepton selections are made for these regions since these variations
relate primarily to jets through parton showers, and not to leptons, and
relaxing cuts can only improve statistical precision.

@@

\paragraph{Alternative $\ttbar$}
Several $\ttbar$ variations model alternative simulations of the
hard scatter (central \powheg\ versus \amcatnlo),
parton shower (central \pythia8 versus \herwig7).
Initial state radiation variations are evaluated from an alternative
sample with $h_\textrm{damp} = 517.5\,\eV[G] \approx 3m(t)$,
which damps real emissions with weights to vary an ``eigentune'' parameter
named \texttt{Var3c}~\cite{ATL-PHYS-PUB-2014-021}.
The central sample uses a default value $h_\textrm{damp} = 172.5\,\eV[G]$ to
approximate $m(t)$~\cite{CMS-PAS-TOP-16-021}.

@@

Scale variations mostly move all regions in the same direction, with increased
scales tending to reduce yields.
Variations in $\mu_R$ are typically around the $10\textrm{--}30\%$ range,
making them larger than the approximately $1\textrm{--}10\%$ variations due to
$\mu_F$.
Given that it takes a maximum, the envelope prescription gives a single larger
variation, but that may not mean more flexibility in the model since the SUSY
prescription has three independent degrees of freedom.

@@

\paragraph{Parton distribution variations}
Like scale variations, uncertainties in parton distribution functions (PDFs)
are assigned by weights stored by the generator.
These weights, however, do not correspond to simple parameter variations,
but different replica PDFs in a Monte Carlo ensemble
that is intended to model the uncertainties left after fitting with the
\textsc{NNPDF} methodology~\cite{NNPDF:2017mvq,BALL2011112,BALL2012608}.

The number of PDF replicas varies between samples, but is between
$100$ and $138$ for all except $\zyjets$, that has just one central and two
alternatives.
Remarkably, both SUSY and PMG prescriptions agree on how to assign variations
from these replica weights --- both say to follow the PDF4LHC
recommendations~\cite{PDF4LHC_Butterworth_2016} (specifically, pp.46,
Section~``6.2. Formulae for the calculation of $PDF$ and $PDF+\alpha_s$
uncertainties).
Since PDF4LHC gives two alternative prescriptions, one using mean and variance
and the other using quantiles, we evaluate and consider both.
Both assignments of PDF variations are negligible in all regions, making
approximately $1\%$ differences before normalization to control regions, but are
included for completeness.
Since there is nothing particularly special about the central PDF weight,
this prescription allows the central scale-factor to differ from $1$,
and this is also a $1\%$-level effect.

@@

\subsection{Signal variations}
\label{sec:2ljets_signal_variations}
Signal uncertainties are a strange case.
Like Standard Model backgrounds, our signal models are simulated by
approximations to their ideal QFT, and those approximations do have
uncertainties that we can describe.
On backgrounds, however, large uncertainties are conservative, since
flexibility to fit outlying data reduces the chance of mistaken claims of
discoveries.
That effect reverses for signals.
In maximum \heplikelihood\ fits, a more flexible
signal model will always fit any observed data better, and in the limit of
infinite flexibility, an unconstrained signal will always fit
any data better than a constrained background, making it look better to certain
interpretations.%
\footnote{%
This superiority of flexible models is not fundamental, but is a property of
best-fit methods.
If we want to punish models that make uncertain predictions, then we could
consider likelihoods (not \emph{maximum} \heplikelihood s), as described in
Section~\ref{sec:searches_searches}.
The important distinction is between uncertain predictions from the model,
and our uncertainty in what the model prediction is.
Our ignorance of a model's predictions should not lead us to excluding it.
}

@@

\subsubsection{Theory variations}
\label{sec:2ljets_theory_variations}
Another signal variation is assigned to cover sources of uncertainty from
scale variations ($\mu_R$, $\mu_F$, and the merging scale
$Q_\textrm{cut}^\textrm{CKKW}$)) and modelling of radiations and parton
showers (`eigentune' parameters
\texttt{Var1},
\texttt{Var2},
\texttt{Var3a},
\texttt{Var3b},
\texttt{Var3c}~\cite{ATL-PHYS-PUB-2014-021}).
Following the \atlas\ SUSY group
prescription~\cite{atlas_twiki_susytheoretical},
these scale variations consider only a coherent movement $\mu_R$ and $\mu_F$
together, so this theory variation includes a total of seven components.

Scale variations now appeared twice for signals:
first for cross-section variations and
second for these theory variations.
They therefore contribute both in each fit model and in the varied
cross-sections for varied exclusion contours.
This is dubious, but thankfully we shall see shortly that this double-counting
of scale variations makes very little difference.

\paragraph{Extrapolation}
To evaluate each variation requires a new simulation with different parameters,
so to reduce costs we use truth-level only
(through SimpleAnalysis~\cite{simpleanalysis_cern}) and evaluate the
variations at only a handful of benchmark signal models.
From these benchmark signal models, we evaluate a relative uncertainty due
due to each component, and extrapolate those variations from these benchmark
examples to others in the parameter plane.

My extrapolation scheme is fairly arbitrary, and is performed as follows.
Each benchmark point $i$ has a location in the parameter plane $\vec y_i$, and
for a given source of uncertainty $j$ its relative variation has been evaluated
as $r_{ij}$.
To extrapolate these variations across each parameter plane to another signal
model at $\vec x$, we use an exponentially weighted combination
\begin{equation}
\label{eqn:2ljets_signal_extrap}
r_j(\vec x) = \frac{
\sum_i r_{ij} e^{-d(\vec x, \vec y_i)}
}{
\sum_i e^{-d(\vec x, \vec y_i)}
}
\end{equation}
in which $d(\vec x, \vec y)$ is a distance function chosen to encode some
physical understanding into the extrapolation,
and the sum is over all benchmark points.

@@

We interpolate an uncertainty for each of the seven components and combine them
all into a single total variation by summation quadrature, capped at a $100\%$
relative error.
This is done independently in each selection region without considering the
directions of coherent variations; all variations are combined as described
and varied together in all regions under the control of a single parameter in
the fit.

\begin{figure}[tp]
\centering
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_signal_sys_c1n2_SRHigh4.pdf}
\caption{C1N2, SR-High-1J}
\end{subfigure}
\hfill
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_signal_sys_c1n2_SROffShell_1.pdf}
\caption{C1N2, SR-OffShell-a}
\end{subfigure}
\\[0.5em]
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_signal_sys_gmsb_SRHigh8_2.pdf}
\caption{GMSB, SR-High-1J}
\end{subfigure}
\hfill
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_signal_sys_gmsb_SRllbb.pdf}
\caption{GMSB, SR-OffShell-a}
\end{subfigure}
\caption[
Example extrapolated signal theoretical uncertainties
]{%
Example extrapolated signal theoretical uncertainties.
Variations are evaluated for signals at the marked points and extrapolated to
each parameter plane according to Equation~\ref{eqn:2ljets_signal_extrap}.
To account for its drastically different kinematics, the C1N2 off-shell space
is treated separately.
}
\label{fig:2ljets_signal_sys_extrap}
\end{figure}

\paragraph{Evaluation}
Statistical precision has limited the utility of most variations described in
this chapter.
Signal variations are no different.
To reduce statistical noise here, I generated truth samples $10$~times as large
as our central, reconstructed samples.
These sample sizes vary roughly with the signal cross-section from $300$k to
$2.8$m events per sample.
Due to the $\sqrt{n}$ scaling of Monte Carlo errors this means approximately
$\sqrt{10}$ times reduced statistical noise.
This remains insufficient, however, since each resulting variation is a
sum in quadrature of variations in $7$ pairs of alternative samples
whose combination will inflate the uncertainty by roughly a factor of
$\sqrt{7}$.
Where signal models' yields has significant in a signal region,
each variation tends to be around $20\%$ relative error,
and all are compatible with the central samples within statistical
uncertainties.
